Hierarchical time series forecasting in Emergency Medical Services

### 2022-11-17 manuscript/03-experiment.qmd
typo and removing pdf

In addition to displaying the trend and seasonality features, we also visualize a few time series at various levels of the aggregation. @fig-dataviz2 reveals different information such as trend, seasonality and noise. For example, some series depict seasonality and trend, whereas some other series report low volume of attended incidents and entropy, making them more volatile and difficult to forecast. At the level on nature of incidents combined with categories of other levels, there are many series that contain zeros with low counts. As such, the data set represents a diverse set of daily time series patterns.

```{r}
#| label: fig-dataviz2
#| cache: true
#| dependson: "fig-feature"
#| out.width: "100%"
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Time series of attended incidents at various levels. The panels show data from the whole country, by control area, by health board, by priority level, and by nature of incident. Only four of the 35 nature of incident categories are shown to avoid too much overplotting."
no_x_axis <- theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank()
)
p_total <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
  autoplot(incident) +
  labs(x = "", y = "Incidents") +
  no_x_axis

p_control <- incident_gthf %>%
  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
  as_tibble() %>%
  select(-nature, -category) %>%
  group_by(date, region) %>%
  summarise(incident = sum(incident), .groups = "drop") %>%
  ggplot(aes(x = date, y = incident, color = factor(region))) +
  geom_line() +
  labs(y = "Incidents", color = "Control") +
  no_x_axis

p_board <- incident_gthf %>%
  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
  as_tibble() %>%
  select(-nature, -category) %>%
  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
  geom_line() +
  # facet_wrap(vars(factor(region)), scales = "free_y") +
  labs(y = "Incidents", color = "Health board") +
  no_x_axis

p_priority <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
  mutate(
    category = recode(category, RED = "Red", AMB = "Amber", GRE = "Green"),
    category = factor(category, levels = c("Red", "Amber", "Green"))
  ) |>
  as_tibble() %>%
  select(-nature, -region) %>%
  ggplot(aes(x = date, y = incident, color = factor(category))) +
  geom_line() +
  scale_color_manual(values = c(Red = "#ff3300", Amber = "#E69f00", Green = "#009e73")) +
  labs(y = "Incidents", color = "Priority") +
  no_x_axis

selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
p_nature <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & !is_aggregated(nature)) %>%
  as_tibble() %>%
  mutate(nature = as.character(nature)) %>%
  filter(nature %in% selected_nature) %>%
  group_by(date, nature, lhb) %>%
  summarise(incident = sum(incident), .groups = "drop") %>%
  ggplot(aes(x = date, y = incident, color = nature)) +
  geom_line() +
  labs(x = "Date", y = "Incidents", color = "Nature of incident")

@@

To evaluate the performance of the various forecasting approaches, we split the data into a series of ten training and test sets. We use a time series cross-validation approach [@hyndman2021forecasting], with a forecast horizon of 84 days, and each training set expanding in 42-day steps. The first training set uses all data up to 2018-04-25, and the first test set uses the 84 days beginning 2018-04-26. The second training set uses all data up to 2018-06-06, with the second test set using the following 84 days. The largest training set ends on 2019-05-09, with the test set ending on 2019-07-31. Model development and hyper-parameter tuning is performed using the training data only. Forecasting performance is evaluated using both point and probabilistic error measures. While we compute forecast errors for the entire 12 weeks, we are most interested in the last 42 days of each test set, because that corresponds to how forecasts are generated for planning in practice. Forecasting performance is evaluated using both point and probabilistic error measures.

>>>>

In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation. @fig-dataviz2 reveals different information such as trend, seasonality and noise. For example, some series depict seasonality and trend, whereas some other series report low volume of attended incidents and entropy, making them more volatile and difficult to forecast. At the level on nature of incidents combined with categories of other levels, there are many series that contain zeros with low counts. As such, the data set represents a diverse set of daily time series patterns.

```{r}
#| label: fig-dataviz2
#| cache: true
#| dependson: "fig-feature"
#| out.width: "100%"
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Time series of attended incidents at various levels. The panels show data from the whole country, by control area, by health board, by priority level, and by nature of incident. Only four of the 35 nature of incident categories are shown to avoid too much overplotting."
no_x_axis <- theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank()
)
p_total <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
  autoplot(incident) +
  labs(x = "", y = "Incidents") +
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few() +
  no_x_axis

p_control <- incident_gthf %>%
  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
  as_tibble() %>%
  select(-nature, -category) %>%
  group_by(date, region) %>%
  summarise(incident = sum(incident), .groups = "drop") %>%
  ggplot(aes(x = date, y = incident, color = factor(region))) +
  geom_line() +
  labs(y = "Incidents", color = "Control")  +
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few() +
  no_x_axis

p_board <- incident_gthf %>%
  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
  as_tibble() %>%
  select(-nature, -category) %>%
  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
  geom_line() +
  # facet_wrap(vars(factor(region)), scales = "free_y") +
  labs(y = "Incidents", color = "Health board") +
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few() +
  no_x_axis

p_priority <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
  mutate(
    category = recode(category, RED = "Red", AMB = "Amber", GRE = "Green"),
    category = factor(category, levels = c("Red", "Amber", "Green"))
  ) |>
  as_tibble() %>%
  select(-nature, -region) %>%
  ggplot(aes(x = date, y = incident, color = factor(category))) +
  geom_line() +
  scale_color_manual(values = c(Red = "#ff3300", Amber = "#E69f00", Green = "#009e73")) +
  labs(y = "Incidents", color = "Priority") +
  ggthemes::theme_few() +
  no_x_axis

selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
p_nature <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & !is_aggregated(nature)) %>%
  as_tibble() %>%
  mutate(nature = as.character(nature)) %>%
  filter(nature %in% selected_nature) %>%
  group_by(date, nature, lhb) %>%
  summarise(incident = sum(incident), .groups = "drop") %>%
  ggplot(aes(x = date, y = incident, color = nature)) +
  geom_line() +
  labs(x = "Date", y = "Incidents", color = "Nature of incident")+
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few()

@@

To evaluate the performance of the various forecasting approaches, we split the data into a series of ten training and test sets. We use a time series cross-validation approach [@hyndman2021forecasting], with a forecast horizon of 84 days, and each training set expanding in 42-day steps. The first training set uses all data up to 2018-04-25, and the first test set uses the 84 days beginning 2018-04-26. The second training set uses all data up to 2018-06-06, with the second test set using the following 84 days. The largest training set ends on 2019-05-09, with the test set ending on 2019-07-31. Model development and hyper-parameter tuning is performed using the training data only. While we compute forecast errors for the entire 12 weeks, we are most interested in the last 42 days of each test set, because that corresponds to how forecasts are generated for planning in practice. Forecasting performance is evaluated using both point and probabilistic error measures.

### 2022-11-17 manuscript/05-results.qmd
typo and removing pdf

# Results and discussion {#sec-result}

@@

@tbl-result-2 also reports point forecast accuracy but using RMSSE. The outperformance of forecast reconciliation using Mint at higher levels of hierarchy is also confirmed by RMSSE. However, for the bottom series, we observe that the base forecasts generated by ETS are the most accurate, followed by its reconciliations forecasts.

>>>>

# Results and discussion {#sec-results}

@@

@tbl-result-2 also reports point forecast accuracy using RMSSE. The outperformance of forecast reconciliation using Mint at higher levels of hierarchy is also confirmed by this measure. However, for the bottom series, we observe that the base forecasts generated by ETS are the most accurate one, followed by its reconciliations forecasts.
